---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo=TRUE}
require(data.table)
```

## Loading and preprocessing the data

```{r, echo=TRUE}
data.raw <- fread(unzip("activity.zip",files="activity.csv"))
head(data.raw)
```

Next, incomplete cases are removed.  
This is accomplished by adding a logical column that's true for complete cases

```{r, echo=TRUE}
data.raw[,comp_case:= complete.cases(steps,date,interval)]
data <- copy( data.raw[comp_case==TRUE] )
head(data)
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
data.ts <- data[,sum(steps),by=date]
names(data.ts) <- c("date","total_steps")
```

Plotting a histogram of the steps

```{r, echo=TRUE}
hist(data.ts$total_steps,main="Daily Step Counts",xlab="Daily Step Count",ylab="Frequency",col="red")
```

Calculating mean

```{r, echo=TRUE}
ts.mean <- mean(data.ts$total_steps)
ts.mean
```

Returning median

```{r, echo=TRUE}
ts.median <- median(data.ts$total_steps)
ts.median
```

## What is the average daily activity pattern?

```{r, echo=TRUE}
data.asbi <- data[,mean(steps),by=interval]
names(data.asbi) <- c("interval","mean_steps")
```

So with step count is averaged for each 5-minute interval, 
plotted to show average levels of activity through the day

```{r, echo=TRUE}
plot(data.asbi$interval,data.asbi$mean_steps,type="l",main="Daily Step Counts",xlab="Interval (minutes)",ylab="Average Step-Count",col="blue")
```

The 5-minute interval with the largest average step count

```{r, echo=TRUE}
maxInterval <- max(data.asbi$mean_steps)
data.asbi[mean_steps==maxInterval]
```

## Inputing missing values

The total missing value can be expressed as the sum of logical columns from the original data table

```{r, echo=TRUE}
with( data.raw, sum(is.na(date))+sum(is.na(interval)) )
with( data.raw, sum(is.na(steps)) )
```

Noting that only steps are missing means all instances of incomplete cases are taged by the 'comp_case' column.  
Missing values are substituted with the average of their date

```{r, echo=TRUE}
averageSteps <- as.integer(mean(data$steps))
data.raw[comp_case==FALSE,steps:= averageSteps]
```

Re-plotting the histogram with updated missing values

```{r, echo=TRUE}
data.ts <- data.raw[,sum(steps),by=date]
names(data.ts) <- c("date","total_steps")
hist(data.ts$total_steps,main="Daily Step Counts (updated)",xlab="Daily Step Count",ylab="Frequency",col="green")
```

## Are there differences in activity patterns between weekdays and weekends?

Adding a factor column to indicate if an entry is for a weekday

```{r, echo=TRUE}
data.raw[,day:= weekdays( as.Date(date, format="%Y-%m-%d") )]
data.raw[,dayType:= ifelse( day=="Sunday" | day=="Saturday", "weekend", "weekday" ) ]
head(data.raw)
```

Making a data set for weekend and weekday data

```{r, echo=TRUE}
# Weekday average steps by interval
data.wkd <- data.raw[dayType=='weekday'][,mean(steps),by=interval]
names(data.wkd) <- c("interval","mean_steps")
# Weekend average steps by interval
data.wke <- data.raw[dayType=='weekend'][,mean(steps),by=interval]
names(data.wke) <- c("interval","mean_steps")
```

Plotting based on average intervals, grouped by dayType

```{r, echo=TRUE}
par(mfrow=c(1, 2), mar=c(4, 4, 1, 1))
# Weekday plot (left)
with(data.wkd, plot( interval, mean_steps, type="l",main="Weekday",xlab="Interval (minutes)",ylab="Average Step-Count",col="blue")  )
# Weekend plot (right)
with(data.wke, plot( interval, mean_steps, type="l",main="Weekend",xlab="Interval (minutes)",ylab="Average Step-Count",col="purple")  )
```






